Using VDMS to Search 100M Images: A Piece of Cake

### 2020-02-12 yfcc100m/paper/vldb2020/eval.tex
Update eval

@@

For all our experiments, we use 2 servers, one hosting VDMS server and
another hosting MySQL server. Both servers have a dual-socket
Intel\textsuperscript{\textregistered}
Xeon\textsuperscript{\textregistered} Platinum 8180 CPU @ 2.50GHz (Skylake),
each CPU with 28 physical cores with multithreading enabled,
for a total of 112 logical cores per server.
The server hosting MySQL has 256GB of DDR4 DRAM, while the server hosting VDMS
has 64GB of DDR4 DRAM. Both servers run Ubuntu 16.04.
The servers and client are connected through a 1GB wired link.

As a baseline, we implemented a similar database system comprised of a
combination of widely available and out-of-the-box components: MySQL Server 5.7,
Apache Web Server 2.4.18, and OpenCV 3.3.
Figure~\ref{fig:systems} shows a logical view of the difference between the
interaction of the client application (which wants to retrieve metadata and
imaged) with VDMS (left) and the baseline (right).
It is worth noting that the images are stored in a shared repository
(ext4 filesystem on a RAID 6 configuration of 16TB) that both Apache WebServer
and VDMS have direct access. In the case of MySQL, metadata is stored in an
attached SSD disk. Even if VDMS has native support for Optane Persistent Memory,
we do not use it in this experiment because of fairness of comparison w.r.t
MySQL. The benefits of Persistent Memory on metadata operations is left
for another paper, and outside the scope of this evaluation.
For this experiment then, we simply use a similar attached SSD disk to store
metadata (even if VDMS Graph store is designed for PM, it can deliver good
performance when using SSDs directly).

@@

We built VDMS and MySQL databases using the YFCC100M dataset with incremental
database sizes 100k, 500k, 1M, 5M, 10M, 50M, and 100M.
For each database size, we created a PMGD graph from the YFCC100m metadata,
the YFCC100m autotags associated with each metadata identifier,
and a list of 1,570 autotags using 100 concurrent VDMS Python clients.
With simple queries, we insert 1,570 nodes for the autotags and insert nodes
for the YFCC media objects with the appropriate metadata.
Adding the autotag confidence scores is a recursive process and requires more
complex queries than adding the previous nodes.
For each metadata identifier in autotags, the query must find the associated
metadata node and tag node.
A connection between the two nodes is created with the confidence score as a
property. The number of metadata nodes are dependent on the database size and
the connections are responsible for 90\% of the elements in each database as
shown in Table~\ref{table:vdmsnodes}.
It is important to note that the metadata identifier, autotags, and
longitude/latitude coordinates are set as indexes of the database to allow
faster retrieval of the metadata.

\begin{table}[h]
\caption{Nodes in VDMS database}
\centering
\begin{tabular}{c c c c}
\hline\hline
Database Size & Connections & Images & Autotags\\
\hline
100k & 848,432      & 100,000     & 1,570\\
500k & 4,249,500    & 500,000     & 1,570\\
1M   & 8,503,045    & 1,000,000   & 1,570\\
5M   & 42,505,478   & 5,000,000   & 1,570\\
10M  & 85,040,404   & 10,000,000  & 1,570\\
50M  & 425,162,070  & 50,000,000  & 1,570\\
100M & 895,572,430  & 99,205,984  & 1,570\\
\hline
\end{tabular}
\label{table:vdmsnodes}
\end{table}

Each MySQL database is created in a similar manner but the data is represented
as three tables: taglist, metadata, and autotags.
By default, MySQL uses infinite number of threads inside InnoDB when processing
requests using four threads per IO read/write.
In some cases, when creating large databases data locks may occur to protect the
data from concurrent updates.
To minimize locking in our runs, we increased the InnoDB buffer pool size to
increase the amount of memory allocated to in-memory data structures
~\cite{mysql,mysql_blog}.
Using a Python client and simple queries, the taglist table is read from the
list of tags with an auto-incremented tagid as an index and the metadata table
is read from the YFCC100m metadata using the identifier as an index.
The autotags table contains the generated autotags and confidence scores for
metadata entries of the metadata table.
To generate the table, we split the autotags data for each database by the
metadata identifier and autotag into new files.
The new files are read into the respective autotags table with the metadata
identifier and tagid as indexes.

\begin{table}[h]
\caption{Rows in MySQL database}
\centering
\begin{tabular}{c c c c}
\hline\hline
 & \multicolumn{3}{c}{Table}\\
\cline{2-4}
Database Size & Autotags & MetaData & Tag List\\
\hline
100k & 848,912     & 100,000    & 1,570\\
500k & 4,241,200   & 498,707    & 1,570\\
1M   & 8,508,380   & 1,000,000  & 1,570\\
5M   & 42,425,905  & 4,987,379  & 1,570\\
10M  & 85,095,265  & 10,000,000 & 1,570\\
50M  & 425,446,208 & 50,000,000 & 1,570\\
100M & 896,002,496 & 99,206,564 & 1,570\\
\hline
\end{tabular}
\label{table:mysqltables}
\end{table}

@@

The VDMS and MySQL databases have comparable number of elements as shown in
Table~\ref{table:vdmsnodes} and ~\ref{table:mysqltables},
but VDMS outperforms MySQL in build speed.
Figure~\ref{fig:db_time_size} illustrates how VDMS can build databases
faster than MySQL as the database size grows.
Key difference in the build times are contributed to the low-level
implementation of how MySQL reads and stores data from the files and the
optimizations (increased InnoDB pool size, etc.) needed
to handle large datasets such as YFCC100m.
On average, it took MySQL 3.72x hours longer to build each database than VDMS.
For example, to build the 100M database, MySQL took 263 hours
while VDMS only needed 72.5 hours.
This is a difference of over 7 days of processing data with VDMS missing less
than 0.001\% images and 0.048\% connections from the entire dataset.

Note that VDMS requires more storage for metadata, shown in
Figure~\ref{fig:db_time_size}.
This is space used to store information about each node/connection.
The Graph Database internal to VDMS (called PMGD) was designed for performance,
specially in environments where persistent memory is present.
This design decision comes as a trade-off for storage footprint, which is
noticeable in our results.
VDMS required 30-41\% more storage than MySQL for storing the same amount
of metadata.
This may become a factor if storage is a limitation, but it should also be noted
that even if we have a 41\% increase in metadata size,
metadata accounts for less than 2\% of the overall database size.
Take, for example, the largest database (both metadata and images) we built
(100M): around 230GB of metadata and 12TB of images.
On the other hand, in systems were persistent memory is a scarce resource,
the increase storage foot print of PMGD may represent a challenge.

\subsection{Images + Metadata}

@@

\begin{figure}[]
\centering
\includegraphics[width=\columnwidth]{figures/concurrency_comparison}
\caption{Concurrency Analysis on Query \"1tag\_resize\"- VDMS vs Baseline}
\label{fig:concurrency_vdms}
\end{figure}

\begin{figure}[t!]
\centering
\includegraphics[width=\columnwidth]{figures/queries_throughput_32}
\caption{Queries - Throughput - 32 Clients}
\label{fig:q_throughput_32}
\end{figure}

@@

An image and metadata search on the YFCC100M databases are very extensive
and would benefit from multithreading to process queries simultaneously.
Figure~\ref{fig:concurrency_vdms} illustrates a concurrency analysis for
the \textit{1tag} query using VDMS to investigate the number of
concurrent clients to use for the per-query performance analysis.
This figure shows the metadata performance plateaus at 32 clients and
the image performance at 56 clients for a simple query.
To further investigate, Figure~\ref{fig:q_throughput_32} and
Figure~\ref{fig:q_throughput_56} illustrate the throughput performance
for the \textit{1tag} and \textit{1tag\_resize} queries for 32 and 56 clients,
respectively.
On the 56-client experiment, a 10x difference occurs between VDMS and
MySQL in metadata throughput while 32 clients is similar but not as high.
The overall performance for both MySQL and VDMS is better with 56 clients;
therefore, we used this value for our per-query analysis.

@@

\subsection{Videos}

@@

\subsection{Feature Vectors}

Another key differentiating factor of VDMS is that it allows the creation of
indexes for high-dimensional feature vectors and the insertion of
these feature vectors associated with entities or visual objects.
Feature vectors are intermediate results of various machine
learning or computer vision algorithms when run on visual data.
Feature vectors are also known as "descriptors" or "visual descriptors".
We use these terms interchangeably.
These descriptors can be labeled and classified to build search indexes,
and there are many in-memory libraries that are designed for
this task~\cite{flann, faiss}.
Using the VDMS API, users can manage feature vector indexes,
query previously inserted elements (images),
run a k-nearest neighbor search (knn) and express relationships
between existing images or descriptors and
the newly inserted descriptors.
By natively supporting descriptors and knn,
VDMS allows out-of-the-box classification
functionalities for many applications.

>>>>

We have used the YFCC100M dataset to evaluate different aspects of our system.
We use the images in the dataset and its associated metadata to implement
an image-search engine based on properties associated to those images.
This is a very common use-case we have encountered when attacking
applications such as smart-retail, sports applications, and video summarization,
where the starting point is a large set of data that must be curated
before proceeding with the data processing and neural network training.
In order to evaluate the different aspects of the performance on the
image search, we built a baseline following the methodology used in the industry, and following what we have done in the past in order to solve
the data search problem, which we describe in Section \ref{images}.

We also cover evaluation we have done using the Video interface VDMS offers,
in order to asses how the implementation performs and scale when used
for video handling.
When it comes to video, ad-hoc solutions have a large number of parameters that
can be tunned, and complex ways to handle the data. Together with that,
there is no system that enables transactional operations over videos files
in the way VDMS does. Because of these two reasons, we focus our efforts in
understanding variations in the the performance of VDMS and its scalability,
rather than comparing it to a baseline that would no represent a fair comparison
for either of the systems. We describe the results in Section \ref{videos}

Similarly, we explore the behaviour of the feature vector functionality in VDMS,
and an evaluation of the different trade-offs that the systems offers for
applications developers. For this, we implemented an image-search application
based on \textit{similarity} search, which we describe in Section \ref{features}

@@

As a baseline, and given that there are no other open-source system
that implements similar functionality and interfaces as VDMS,
we implemented an equivalent visual data management system comprised of a
combination of widely available, off-the-shelve components:
MySQL Server 5.7 (for storing metadata),
Apache Web Server 2.4.18 (as interface for image access), and
OpenCV 3.3 (to provide pre-processing operations on images).
We have implemented a set of client-side applications that takes care
of retrieving the components from the different systems, and applies
pre-processing operations when needed.
For visual data workloads, building an approach like the one
implemented as a baseline for this work
is the common practice in the industry \cite{heystack, tao}.

For all our experiments, we use 2 servers, one hosting a VDMS server and
another hosting a the Baseline implementation.
Both servers have a dual-socket Intel\textsuperscript{\textregistered}
Xeon\textsuperscript{\textregistered} Platinum 8180 CPU @ 2.50GHz (Skylake),
each CPU with 28 physical cores with multithreading enabled,
for a total of 112 logical cores per server.
The server hosting MySQL has 256GB of DDR4 DRAM, while the server hosting VDMS
has 64GB of DDR4 DRAM. Both servers run Ubuntu 16.04.
The client application (running the queries and measuring round-trip time)
is connected to the server through a 1GB wired link through
a 10GB-backplane switch.

Figure~\ref{fig:systems} shows a logical view of the difference between the
interaction of the client application (which wants to retrieve metadata and
imaged) with VDMS (left) and the baseline (right).
The client application was implemented using Python 3 for both VDMS and
the baseline.

It is worth noting that the images are stored in a shared repository
(ext4 filesystem on a RAID 6 configuration of 16TB) that both Apache WebServer
and VDMS have direct access. In the case of the baseline, metadata is
stored in MySQL using an attached SSD disk.
Even if VDMS has native support for Optane Persistent Memory,
we do not use it in this experiment because of fairness of comparison w.r.t
MySQL (which was not design for Persistent Memory type of storage).
The benefits of Persistent Memory on metadata operations is left
for another paper, and outside the scope of this evaluation.
For this experiment then, in the case of VDMS we simply use a similar
attached SSD disk to store metadata.
Even if PMGD, the graph database used by VDMS, is designed for PM,
it can deliver good performance when using SSDs directly.

@@

For the metadata, we built VDMS and MySQL databases
using the YFCC100M dataset with incremental database sizes.
For simplicity, we named the database based on the number of images
it contains, as follows: 100k, 500k, 1M, 5M, 10M, 50M, 100M.
The exact number of images/elements on each database is shown in
Table~\ref{table:vdmsnodes} and ~\ref{table:mysqltables}
For each database size, we created a PMGD graph from the YFCC100m metadata,
the YFCC100m autotags associated with each metadata identifier,
and the list of 1,570 autotags using 100 concurrent VDMS Python clients.
With simple queries, we insert 1,570 nodes for the autotags and insert nodes
for the YFCC media objects with the appropriate metadata.
Adding the autotag confidence scores is a recursive process and requires more
complex queries than adding the previous nodes.
For each metadata identifier in autotags, the query must find the associated
metadata node and tag node.
A connection between the two nodes is created with the confidence score as a
property. The number of metadata nodes are dependent on the database size and
the connections are responsible for 90\% of the elements in each database as
shown in Table~\ref{table:vdmsnodes}.
It is important to note that the metadata identifier, autotags, and
longitude/latitude coordinates are set as indexes of the database to allow
faster retrieval of the metadata.

\begin{table}[h]
\caption{VDMS Database - Number of Elements}
\centering
\begin{tabular}{c c c c}
\hline\hline
Database Name & \# of Images & \# of Connections & \# of Autotags\\
\hline
100k & 100,000     & 848,432      & 1,570\\
500k & 500,000     & 4,249,500    & 1,570\\
1M   & 1,000,000   & 8,503,045    & 1,570\\
5M   & 5,000,000   & 42,505,478   & 1,570\\
10M  & 10,000,000  & 85,040,404   & 1,570\\
50M  & 50,000,000  & 425,162,070  & 1,570\\
100M & 99,205,984  & 895,572,430  & 1,570\\
\hline
\end{tabular}
\label{table:vdmsnodes}
\end{table}

Each MySQL database is created in a similar manner but the data is represented
as three tables, following the relational model:
1) \textit{metadata} table, contains one row per image,
and a column for each property
associated with the images (some of which are listed in \ref{dataset});
2) \textit{taglist} table: contains one row per each autotag elements
(always 1570 rows);
3) \textit{autotags} table: contains on row per each autotag
assigned to an image, each row containing a foreign key to the
image and a foreign key to the tag, and
the probability assigned to that tag belonging to that image.
Given that there are 8 autotags, on average, per image, the \textit{autotags}
table has around 8 times the number of rows present in the
\textit{metadata} table, as can be seen in~\ref{table:mysqltables}.

THIS NEEDS FIX, NOT SURE I UNDERSTAND WHAT YOU MEAN:
By default, MySQL uses infinite number of threads inside InnoDB when processing
requests using four threads per IO read/write.
In some cases, when creating large databases data locks may occur to protect the
data from concurrent updates.
To minimize locking in our runs, we increased the InnoDB buffer pool size to
increase the amount of memory allocated to in-memory data structures
~\cite{mysql,mysql_blog}.

Using a Python client and simple queries, the taglist table is read from the
list of tags with an auto-incremented tagid as an index and the metadata table
is read from the YFCC100m metadata using the identifier as an index.
The autotags table contains the generated autotags and confidence scores for
metadata entries of the metadata table.

To generate the table, we split the autotags data for each database by the
metadata identifier and autotag into new files.
The new files are read into the respective autotags table with the metadata
identifier and tagid as indexes.

\begin{table}[h]
\caption{MySQL Database - Number of Elements}
\centering
\begin{tabular}{c c c c}
\hline\hline
 & \multicolumn{3}{c}{Table}\\
\cline{2-4}
Database Name & \# of Images & \# of Image Tags & Tag List\\
\hline
100k & 100,000    & 848,912     & 1,570\\
500k & 498,707    & 4,241,200   & 1,570\\
1M   & 1,000,000  & 8,508,380   & 1,570\\
5M   & 4,987,379  & 42,425,905  & 1,570\\
10M  & 10,000,000 & 85,095,265  & 1,570\\
50M  & 50,000,000 & 425,446,208 & 1,570\\
100M & 99,206,564 & 896,002,496 & 1,570\\
\hline
\end{tabular}
\label{table:mysqltables}
\end{table}

@@

The VDMS and MySQL databases have comparable number of elements as shown in
Table~\ref{table:vdmsnodes} and ~\ref{table:mysqltables}, and the differences
can be attributed to failures in data preparation/loading because of
incomplete/inconsistent formating, which is common on large datasets.
The difference is less than 0.1\% in terms of number of elements (images and/or
metadata information).

\subsubsection{Database Loading Time}

One of the first things we noted is the difference is loading times,
where VDMS outperforms MySQL by a large margin.
Figure~\ref{fig:db_time_size} illustrates how VDMS can build databases
faster than MySQL, and how the speedup is sustained as the database size grows.
Key difference in the build times are attributed to the low-level
implementation of how MySQL reads and stores data from the files and the
optimizations (increased InnoDB pool size, etc.) needed
to handle large datasets such as YFCC100m.
On average, it took MySQL 3.72x longer to build each database than VDMS.
For example, to build the 100M database, MySQL took 263 hours
while VDMS only needed 72.5 hours.

\subsubsection{Database Storage Footprint}

Another aspect important to note is that
VDMS requires more storage for metadata, shown in Figure~\ref{fig:db_time_size}.
This is space used to store information about each node/connection.
The Graph Database internal to VDMS (called PMGD) was designed for performance,
specially in environments where persistent memory is present.
This design decision comes as a trade-off for storage footprint, which is
noticeable in our results.
VDMS required 30-41\% more storage than MySQL for storing the same amount
of metadata.
This may become a factor if storage is a limitation, but it should also be noted
that even if we have a 41\% increase in metadata size,
metadata accounts for less than 2\% of the overall database size.
Take, for example, the largest database (both metadata and images) we built
(100M): around 230GB of metadata and 12TB of images.
On the other hand, in systems were persistent memory is a scarce resource,
the increase storage foot print of PMGD may represent a challenge.

%=========================================

\subsection{Images + Metadata}
\label{images}

@@

\begin{figure}[]
\centering
\includegraphics[width=\columnwidth]{figures/concurrency_comparison}
\caption{Concurrency Analysis on Query "1tag\_resize\- VDMS vs Baseline}
\label{fig:concurrency_vdms}
\end{figure}

% We decided to remove this image because it is redudant with the concurrency
% analysis.
% \begin{figure}[t!]
% \centering
% \includegraphics[width=\columnwidth]{figures/queries_throughput_32}
% \caption{Queries - Throughput - 32 Clients}
% \label{fig:q_throughput_32}
% \end{figure}

@@

An image and metadata search on the YFCC100M databases are very extensive
and would benefit from multithreading to process queries simultaneously.
Figure~\ref{fig:concurrency_comparison} illustrates a concurrency analysis for
the \textit{1tag} query using VDMS to investigate the number of
concurrent clients to use for the per-query performance analysis.
This figure shows the metadata performance plateaus at 32 clients and
the image performance at 56 clients for a simple query.
% To further investigate, Figure~\ref{fig:q_throughput_32} and
% Figure~\ref{fig:q_throughput_56} illustrate the throughput performance
% for the \textit{1tag} and \textit{1tag\_resize} queries for 32 and 56 clients,
% respectively.
On the 56-client experiment, a 10x difference occurs between VDMS and
MySQL in metadata throughput while 32 clients is similar but not as high.
The overall performance for both MySQL and VDMS is better with 56 clients;
therefore, we used this value for our per-query analysis.

@@

%=========================================

\subsection{Videos}
\label{videos}

@@

%=========================================

\subsection{Feature Vectors}
\label{features}

Another key differentiating factor of VDMS is that it allows the creation of
indexes for high-dimensional feature vectors and the insertion of
these feature vectors associated with entities or visual objects.
Feature vectors are intermediate results of various machine
learning or computer vision algorithms when run on visual data.
Feature vectors are also known as \textit{descriptors}
or \textit{visual descriptors}. We use these terms interchangeably.
These descriptors can be classified and labeled and used to build search indexes,
and there are many in-memory libraries that are designed for
this task~\cite{flann, faiss}.
Using the VDMS API, users can manage feature vector indexes,
query previously inserted elements (images),
run a k-nearest neighbor search (knn), and express relationships
between existing images or descriptors and
the newly inserted descriptors.
By natively supporting descriptors and knn,
VDMS allows out-of-the-box classification functionalities for many applications.
